---
title: "Blackpoll_Warbles"
author: "Joel"
date: "2/25/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
require(tidyverse)
```

#LOAD IN NECESSARY PACKAGES
```{r, echo=TRUE}
library(tidyverse)
library(lubridate)
library(reshape2)
```

#VISUALIZE THE DATA FRAME
```{r}
str(blpw.all$band)
blpw.all$band <- factor(blpw.all$band)
table(blpw.all$band)
```  

# FIND THOSE BIRDS WITH BAND NUMBER RECORDED MORE THAN ONCE
```{r}
# this identifies the number of duplcated band numbers
recaptured <- anyDuplicated(blpw.all$band, fromLast = TRUE)
# this subsets thoes duplications
recap <-  blpw.all %>% group_by(band, location) %>% filter(n()>1) 
      # all the ones recaptured DO have an R in the recapture column
```  

# SEPARATE THE DATA ACCORDING TO THE TWO LOCATIONS
```{r}

table(recap$location) # good there are only two locations
recap <- ungroup(recap)
    # select thoes on Seal Island
seal <- filter(recap, location == "ABO-SI")
# select thoes on BP Island
bp <- filter(recap, location == "ABO-BP")
# so there are only 38 observations
# there was 40 observations when filtering by recapture column
```


#---- FIND THE DIFFERENCE IN WEIGHT UPON RECAPTURE AT SEAL ISLAND --------------------
```{r}
seal <- select (seal, location, band, mass, year, month, day)

# creates a date column so we can organize the captures by date
seal$date <- with(seal, ymd(sprintf('%04d%02d%02d', year, month, day)))
```
# now that we seperated Seal from BP, there will be a few single-bird observations,
# indecisive bird removal:
```{r}
sealrecaptured <- anyDuplicated(dogfood$band, fromLast = TRUE)
sealrecap <-  seal %>% group_by(band, year) %>% filter(n()>1) 
```

#GROUP BY BAND NUMBER SO WE CAN FIND DIFFERENCE IN MASS UPON RECAPTURE
```{r}
dogfood <- group_by(sealrecap, band, year)
```

#SELECTING BIRDS GROUPED BY BAND OF THE EARLIER DATE OF CAPTURE
```{r}
topn <- top_n(x = dogfood, n = -1, wt = date)
dogfood <- ungroup(dogfood)
topn <- ungroup(topn)
```
# RENAMING NEW WEIGHTS COLUMN "mass" TO "firstmass"
```{r}
colnames(topn)[3] <- "firstmass"
```
#JOINING THE TABLES TOGETEHER, WHILE DUPLICATING "firstmass"
```{r}
brezhnev <- left_join(dogfood, topn, by = "band") # proposed changing brezhnev to antifreeze
```
# MUTATING A NEW COLUMN OF BIRDS WITH SASS (MASS-INITIAL = SASS)
```{r}
brezhnev <- ungroup(brezhnev)
brezhnev <- group_by(brezhnev, band) %>%
  mutate(sass = mass-firstmass) %>%
  mutate(yday = (yday(date.x))) # dont need this whole yday thing here
```

# Try graphing this and see what kind of magic happens
```{r}
ggplot(data = brezhnev,mapping = aes(x = yday, y = sass, colour = band), show.legend = FALSE) +
  xlab("Time of Year")+ylab("Mass (in grams, relative to capture date)") +
  geom_point(show.legend = FALSE) +
  geom_line(show.legend = FALSE) +
  theme_bw()
```


#---- FIND THE DIFFERENCE IN WEIGHT UPON RECAPTURE AT BON PORTAGE--------------------
```{r}
bp <- select (bp, location, band, mass, year, month, day)
bp$date <- with(bp, ymd(sprintf('%04d%02d%02d', year, month, day))) 

# Manage them indecisive birds:
bprecaptured <- anyDuplicated(grumpy_old_man$band, fromLast = TRUE)
bprecap <-  bp %>% group_by(band, year) %>% filter(n()>1)
```

#GROUP BY BAND NUMBER SO WE CAN FIND DIFFERENCE IN MASS UPON RECAPTURE
```{r}
grumpy_old_man <- group_by(bprecap, band, year) #%>%
#arrange(desc(date))
# dont need the arrange really

# topline group - (finding initial mass)
tops <- top_n(x = grumpy_old_man, n = -1, wt = date)
grumpy_old_man <- ungroup(grumpy_old_man)
tops <- ungroup(tops)

# renamed column "mass" to "firstmass"
colnames(tops)[3] <- "firstmass"

# joining for the mutate calculation
Bertha <- left_join(grumpy_old_man, tops, by = "band")

# calculating sassy-ness
Bertha <- ungroup(Bertha)
Bertha <- group_by(Bertha, band) %>%
  mutate(sass = mass-firstmass) %>%
  mutate(yday = (yday(date.x)))
```

#---- Try graphing this and see what kind of magic happens-------------------------
```{r}
ggplot(data = Bertha,mapping = aes(x = yday, y = sass, colour = band), show.legend = FALSE) +
  xlab("Time of Year")+ylab("Mass (in grams, relative to capture date)") +
  geom_point(show.legend = FALSE) +
  geom_line(show.legend = FALSE) +
  theme_bw()
# magical...
```






