---
title: "Make the Plot Assignment"
author: "Joel Goodwin & Mike Light"
date: "2/24/2018"
output: html_document
---
#Load in Data
```{r, echo=FALSE}
setwd("/Users/joelgoodwin/Google Drive/School/Masters/Classes/Biol 5023- Research Methods II/VarroaETG_Percent/Biol5023/BlackpollAssignment")

blpw.all<- readRDS("blpw.all.RDS")
blpw.all <- blpw.all
```

```{r, echo=TRUE, eval=FALSE}
blpw.all<- readRDS("blpw.all.RDS")
blpw.all <- blpw.all
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
require(tidyverse)
```

#Load in Necessary Packages
```{r, echo=TRUE}
library(tidyverse)
library(lubridate)
library(reshape2)
library(ggplot2)
library(dplyr)
```

#Begin by Visualizing the Data Frame
```{r, echo=TRUE}
str(blpw.all)
blpw.all$band <- factor(blpw.all$band)
```  

#Select Those Birds Whose Band Number Was Recorded More Than Once
```{r, echo=TRUE}
# this subsets duplications
recap <-  blpw.all %>% group_by(band, location) %>% filter(n()>1)
# all the recaptured birds DO have an R in the recapture column
```  

#Separate the Data According to Location For Ease of Analysis
We want to group the data by location and separate this into two separate data frames to ease our analysis. One data frame for Bon Portage and one for Seal Island.
```{r, echo=TRUE}
table(recap$location) 
```

#Find the difference in Weight Upon Recapture at Seal Island 
###Select those birds on Seal Island
```{r, echo=TRUE}
seal <- filter(recap, location == "ABO-SI")
```

```{r, echo=TRUE}
seal <- select (seal, location, band, mass, year, month, day)

# The following code creates a date column so we can organize the captures by date
seal$date <- with(seal, ymd(sprintf('%04d%02d%02d', year, month, day)))
```

###Indecisive bird removal
Now that we have separated Seal Island observations from Bon Portage Island observations, there will be a few single-bird observations. We remove them from analysis.
```{r}
sealrecaptured <- anyDuplicated(seal$band, fromLast = TRUE)
sealrecap <-  seal %>% group_by(band, year) %>% filter(n()>1)
```

###Group by Band Number so we can Find Difference in Mass Upon Recapture
```{r}
sealrecap <- group_by(sealrecap, band, year)
```

###Selecting Birds Grouped by Band of the Earlier Date of Capture
```{r}
topn <- top_n(x = sealrecap, n = -1, wt = date)
sealrecap <- ungroup(sealrecap)
topn <- ungroup(topn)
```

###Rename new weights column to "firstmass"
```{r}
colnames(topn)[3] <- "firstmass"
```

###Joining the Tables Together while duplicating "firstmass"
```{r}
joined_table <- left_join(sealrecap, topn, by = "band") 
```
###Mutating a New Column of Birds with Difference in Mass (Mass-Initial Mass = SASS)
```{r}
joined_table <- ungroup(joined_table)
joined_table <- group_by(joined_table, band) %>%
  mutate(sass = mass-firstmass)
```

#Find the Difference in Weight Upon Recapture at Bon Portage
###Select those birds on Bon Portage Island
```{r}
bp <- filter(recap, location == "ABO-BP")
```

```{r}
bp <- select (bp, location, band, mass, year, month, day)
bp$date <- with(bp, ymd(sprintf('%04d%02d%02d', year, month, day))) 

bprecap <-  bp %>% group_by(band, year) %>% filter(n()>1)
```
###Indecisive bird removal
Now that we have separated Bon Portage observations from Seal Island observations, there will be a few single-bird observations. We remove them from analysis.
```{r}
bprecaptured <- anyDuplicated(bprecap$band, fromLast = TRUE)
```

###Group by Band Number so we can Find Difference in Mass Upon Recapture
```{r}
bprecap<- group_by(bprecap, band, year) 
```
###Selecting Birds Grouped by Band of the Earlier Date of Capture
```{r}
tops <- top_n(x = bprecap, n = -1, wt = date)
bprecap <- ungroup(bprecap)
tops <- ungroup(tops)
```
###Rename new weights column from "mass" to "firstmass"
```{r}
colnames(tops)[3] <- "firstmass"
```

###Joining the Tables Together While Duplicating "firstmass"
```{r}
joined_table_bp <- left_join(bprecap, tops, by = "band")
```
###Mutating a New Column of Birds with Difference in Mass (Mass-Initial Mass = SASS)
```{r}
joined_table_bp <- ungroup(joined_table_bp)
joined_table_bp <- group_by(joined_table_bp, band) %>%
  mutate(sass = mass-firstmass)
```

#Re-Joining the Data From Both Islands
```{r}
# now to combine BP with SEAL
ungroup(joined_table_bp, joined_table)
jointest <- full_join(joined_table_bp, joined_table)
jointest <- data.frame(jointest)
```

### Create a month-day column 
```{r}
monoyear <- rep("2000",times = 259) 
# Makes a column with just the year 2000 (arbitrarily selected year)

jointest <- cbind(monoyear, jointest) 
# Add monoyear (2000) to jointest

# Now we must change year 2000 to numeric
jointest$monoyear <- as.numeric(as.character(jointest$monoyear))
```

###Make All Dates from a Single Year to Make the Graph Work
```{r}
jointest <- jointest %>%
  mutate(monodate = make_date(monoyear,month.x, day.x))

```

#The Fruit of Our Labour
After many dead ends, we've managed to create the graph we've been aiming for.
```{r, echo = TRUE}
ggplot(data = jointest,mapping = aes(x = monodate, y = sass, colour = band), show.legend = FALSE) +
  xlab("Time of Year") + 
  ylab("Mass (in grams, relative to capture date)") +
  geom_point(group= jointest$band, show.legend = FALSE) +
  geom_line(group= jointest$band, show.legend = FALSE) +
  facet_wrap(~location.x) +
  theme_gray()
```
